### 2024-03-22 camera-ready-submission/source/main.tex

\pdfminorversion=7
% Options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[sigconf, screen]{acmart}

\usepackage{xcolor}
%\usepackage{cleveref}
\usepackage{color}
\usepackage{fancyvrb}
\usepackage{acmart-taps}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

%\ifLuaTeX
%  \usepackage{selnolig}  % disable illegal ligatures
%\fi
%\usepackage[]{natbib}
%\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
%\urlstyle{same} % disable monospaced font for URLs

\widowpenalty=10000
\clubpenalty=10000
\displaywidowpenalty=10000

\copyrightyear{2023}
\acmYear{2023}
\setcopyright{rightsretained}
\acmConference[IFL 2023]{The 35th Symposium on Implementation and Application of Functional Languages}{August 29--31, 2023}{Braga, Portugal}
\acmBooktitle{The 35th Symposium on Implementation and Application of Functional Languages (IFL 2023), August 29--31, 2023, Braga, Portugal}
\acmDOI{10.1145/3652561.3652565}
\acmISBN{979-8-4007-1631-7/23/08}
\keywords{dependent types, elaborators, elaboration, open datatype, extensible datatype, type-checking}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
       <concept_desc>Software and its engineering~Functional languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011074.10011075.10011077</concept_id>
       <concept_desc>Software and its engineering~Software design engineering</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003790.10011740</concept_id>
       <concept_desc>Theory of computation~Type theory</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10002944.10011123.10011673</concept_id>
       <concept_desc>General and reference~Design</concept_desc>
       <concept_significance>100</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Functional languages}
\ccsdesc[300]{Software and its engineering~Software design engineering}
\ccsdesc[300]{Theory of computation~Type theory}
\ccsdesc[100]{General and reference~Design}

\begin{document}

\title{ExEl: Building an Elaborator Using Extensible Constraints}

\author{Bohdan Liesnikov}
\orcid{0009-0000-2216-8830}
\affiliation{
  \institution{Delft University of Technology}
  \city{Delft}
  \country{The Netherlands}}
\email{B.Liesnikov@tudelft.nl}
\author{Jesper Cockx}
\orcid{0000-0003-3862-4073}
\affiliation{
  \institution{Delft University of Technology}
  \city{Delft}
  \country{The Netherlands}}
\email{J.G.H.Cockx@tudelft.nl}

\hypersetup{
  pdftitle={ExEl: Building an Elaborator Using Extensible Constraints},
  pdfauthor={Bohdan Liesnikov and Jesper Cockx},
  pdfkeywords={dependent types, elaborators, elaboration, open datatype, extensible datatype, type-checking},
  pdfcreator={LaTeX via pandoc},
  pdfproducer={pdftex},
  hidelinks,
}

\begin{abstract}
  Proof assistants and dependently typed languages such as Coq, Agda, Lean, and Idris can be used to ascertain the correctness of software with mathematical precision.
  While much research has been done on their theoretical foundations, their actual implementations have been studied to a much lesser extent.
  As a result, features that are not considered part of the theoretical foundations - such as implicit arguments and type classes - have their own bespoke implementation for each language, making for code bases that are hard to understand and maintain.
  To address some of the common problems in the implementations of dependently typed languages, we present a modular architecture for implementing the transformation from user-friendly surface syntax into a small and well-behaved core language, also known as elaboration.
  Our architecture is made modular through the use of an open datatype of constraints and a plugin system for solvers that work on these constraints, which means that each new feature is contained in its own module.
  We showcase our design with a proof-of-concept elaborator for a language with dependent types, implicit arguments, higher-order unification, and instance arguments.
\end{abstract}

\maketitle

\input{body.tex}

\begin{acks}
  Jesper Cockx holds an \grantsponsor{nwoviveni202216}{NWO Veni}{https://www.nwo.nl/en/projects/viveni202216} grant on `A trustworthy and extensible core language for Agda' (\href{https://www.nwo.nl/en/projects/viveni202216}{VI.Veni.202.216}).
\end{acks}

\bibliographystyle{ACM-Reference-Format}

\bibliography{bib.bib}

\end{document}